---
title: "p8105_hw2_ww2745"
author: "ww2745"
date: "2024-10-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1
Read and clean the data.

```{r nyc_data}
nyc_tran_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada
    ) |>  
  mutate(
    entry = case_match(entry,
                            "YES" ~ TRUE,
                            "NO" ~ FALSE),
    vending = case_match(vending,
                            "YES" ~ TRUE,
                            "NO" ~ FALSE))



  
print(nyc_tran_df, n = 12)

```

The dataset contains the following variables: `r colnames(nyc_tran_df)`. For the previous chunck, we have loaded the NYC transit dataset, clean the variable names, select the columns we desire and change the type of the variable 'entry'. There are `r nrow(nyc_tran_df)` rows and `r ncol(nyc_tran_df)` columns in the `nyc_tran_df`. These data are not tidy.


```{r answer_question, include = FALSE}
station_distinct =
  distinct(
    nyc_tran_df, station_name, line, .keep_all = TRUE
    )

ada_compliant = 
  filter(
    station_distinct, ada == TRUE
  )

```

```{r answer_question_prop, include = FALSE}
a = 
  filter(
    nyc_tran_df, vending == FALSE & entry == TRUE
  ) |> 
  nrow()

b = 
  filter(
    nyc_tran_df, vending == FALSE
  ) |>  
  nrow()

```

+ There are `r nrow(station_distinct)` distinct stations in NYC.

+ `r nrow(ada_compliant)` of them are ADA compliant.

+ The proportion of station entrances / exits without vending allow entrance is `r a/b`.


Reformat the data.
```{r reformat_data}

dist_route_df = station_distinct |> 

mutate_at(vars(route8:route11), as.character) |> 
pivot_longer(
  route1:route11,
  names_to = "route",
  values_to = "train")

head(dist_route_df)
```

 + There are `r nrow(filter(dist_route_df, train == "A"))` distinct stations serve the A train.
 
 + There are `r nrow(filter(dist_route_df, train == "A" & ada == "TRUE"))` distinct stations serve the A train and are ADA compliant.

# Problem 2

Read and create `mr_trash` dataset.
```{r import_data, warning=FALSE}
mr_trash = 
  read_excel(
  "./data/202409 Trash Wheel Collection Data.xlsx",
  sheet = "Mr. Trash Wheel",
  range = ("A2:N655")) |> 
  janitor::clean_names() |>  
  drop_na(dumpster) |>  
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) |> 
  mutate(trash_wheel = "Mr. Trash Wheel") 
```

Read and clean organize the data for Professor Trash Wheel and Gwynnda.
```{r prof_and_gwy_data}
prof_df =
  read_excel(
    "./data/202409 Trash Wheel Collection Data.xlsx",
		sheet = "Professor Trash Wheel",
		range=("A2:M123")	) |> 
	janitor::clean_names() |> 
	drop_na(dumpster) |> 
  mutate(trash_wheel = "Professor Trash Wheel") 


gwy_df = 
	read_excel(
		"./data/202409 Trash Wheel Collection Data.xlsx",
		sheet = "Gwynnda Trash Wheel",
		range=("A2:L266")	)|> 
	janitor::clean_names() |> 
	drop_na(dumpster) |> 
  mutate(trash_wheel = "Gwynnda Trash Wheel")
```

Combine three datasets.
```{r merge_data}
combined_data =
  left_join(prof_df, gwy_df, by = "trash_wheel")  
combined_data =
  left_join(combined_data, mr_trash, by = "trash_wheel") 

```

This dataset contains information from the Mr. Trashwheel trash collector,Professor Trash Wheel and Gwynnda. The dataset contains information on year, month, and trash collected, include some specific types of trash. There are a total of `r nrow(combined_data)` rows in the final dataset. 


  + The total weight of trash collected by Professor Trash Wheel was `r mr_trash |>  pull(weight_tons) |> sum()` tons.
  
  + The total number of cigarette butts collected by Gwynnda in June of 2022 was `r gwy_df |>  filter(year == 2022 & month == "June",)  |>  pull(cigarette_butts) |>  sum()`.

